---
title: "Vector Autoregressive (VAR)"
output: github_document
date: "2024-03-18"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

#### Libraries 

```{r}
library(vars)
library(tsDyn)
library(tseries)
library(stargazer)

```

#### Simalating bivariate VAR(2)

```{r}
set.seed(111)
A1 <- matrix(c(0.5, 0.4, 0.1, 0.5), 2)
A2 <- matrix(c(0, 0.25, 0, 0), 2)
B <- cbind(A1, A2)
var <- VAR.sim(B = B, n = 250, include = "none", lag = 2)
var_ts <- data.frame(y1_t = var[, 1], y2_t = var[, 2])
plot.ts(var_ts)

```

#### Checking stationarity of series

```{r}
adf.test(var_ts$y1_t)
adf.test(var_ts$y2_t)

```

#### fitting a VAR model to simalated data

```{r}
lag <- VARselect(var_ts, lag.max = 5)
lag
model <- VAR(var_ts, p = 2, type = "non")
stargazer(model$varresult, type = "text", style = "aer", title = "VAR(2)", columns.label = c("y1_t", "y2_t"))

```
#### Checking the stability

```{r}
roots(model, modulus = TRUE)

```

#### CUMSUM test and serial correlation

```{r}
stability_model <- stability(model)
plot(stability_model$stability$y1_t, lwd = 2)
plot(stability_model$stability$y2_t, lwd = 2)
sctest(model)
serial <- serial.test(model, lags.pt = 8)
plot(serial, names = "y1_t")
plot(serial, names = "y2_t")

```

#### Test for heteroscedasticity  
```{r}
arch.test(model, lags.multi = 12, multivariate.only = TRUE)

```
##### The p-value above 0.05 suggests homoscedasticity, or constant variance.

#### Normality test

```{r}
normality.test(model, multivariate.only = TRUE)

```
##### The p-value suggests that the normality assumption for the residuals is met.


#### Granger and instantaneous causality

```{r}
granger_y1 <- causality(model, cause = "y1_t")
granger_y2 <- causality(model, cause = "y2_t")

granger_y1
granger_y2

```

#### Impuse Responce 

```{r}
irf11 <- irf(model, impulse = "y1_t", response = "y1_t", n.ahead = 40, 
boot = TRUE, run = 100, ci = 0.95)
plot(irf11)

irf12 <- irf(model, impulse = "y1_t", response = "y2_t", n.ahead = 40, 
boot = TRUE, run = 100, ci = 0.95)
plot(irf12)


irf22 <- irf(model, impulse = "y2_t", response = "y2_t", n.ahead = 40, 
boot = TRUE, run = 100, ci = 0.95)
plot(irf22)


irf21 <- irf(model, impulse = "y2_t", response = "y1_t", n.ahead = 40, 
boot = TRUE, run = 100, ci = 0.95)
plot(irf21)

```


##### Here we note that a positive shock to y1_t increases y2_t by a relatively large and persistent amount.



#### Forecast Error Variance decomposition (FEVD)

```{r}
vd <- fevd(model, n.ahead = 15)
plot(vd)
```

#### Forecast

```{r}
pred <- predict(model, n.ahead = 20)
plot(pred, names = "y1_t")
plot(pred, names = "y2_t")

fanchart(pred, names = "y1_t", color = c("skyblue4", "skyblue3", "skyblue2", "skyblue1", 
"skyblue", "purple4", "purple3", "purple2", "purple1", col.y = "skyblue"))

fanchart(pred, names = "y2_t", color = c("skyblue4", "skyblue3", "skyblue2", "skyblue1", 
"skyblue", "purple4", "purple3", "purple2", "purple1", col.y = "skyblue"))

```




